\thornado-hydro: towards discontinuous galerkin methods for supernova hydrodynamics

### 2018-11-13 Documents/Euler/Astronum_2018/ms_endeve.tex
Addition of refs.  Already in submitted version

\newcommand{\deriv}[2]{\frac{d #1}{d #2}}
\newcommand{\pderiv}[2]{\frac{\partial #1}{\partial #2}}
\newcommand{\vect}[1]{\boldsymbol{#1}}
\newcommand{\f}[2]{\frac{#1}{#2}}
\newcommand{\dx}{\Delta x}
\newcommand{\pd}[2]{\partial_{#2}{#1}}
\newcommand{\bx}{\vect{x}}
\newcommand{\bK}{\vect{K}}
\newcommand{\cT}{\mathcal{T}}
\newcommand{\xL}{x_{\textnormal{\tiny\textsc{L}}}}
\newcommand{\xH}{x_{\textnormal{\tiny\textsc{H}}}}
\newcommand{\xMin}{x_{\textnormal{\tiny m}}}
\newcommand{\xMax}{x_{\textnormal{\tiny M}}}
\newcommand{\trans}{\textnormal{\tiny\textsc{T}}}
\newcommand{\leftState}{\textnormal{\tiny\textsc{L}}}
\newcommand{\rightState}{\textnormal{\tiny\textsc{R}}}
\newcommand{\TCI}{\textnormal{\tiny\textsc{TCI}}}
\newcommand{\shock}{\textnormal{\tiny\textsc{Sh}}}
\newcommand{\thornado}{\texttt{thornado}}

@@

We have presented preliminary algorithm details and numerical results for solvers of the non-relativistic and special relativistic Euler equations of gas dynamics as implemented in the {\bf t}oolkit for {\bf h}igh-{\bf or}der {\bf n}eutrino-r{\bf ad}iation hydr{\bf o}dynamics (\thornado).  
The spatial discretization is based on the DG method and the ODEs resulting from this discretization are integrated in time using SSP-RK methods.  
We employ a spectral-type nodal collocation DG approximation based on Legendre-Gauss points for interpolation and numerical quadrature evaluation \cite{bassi_etal_2013}.  
This choice simplifies expressions for the semi-discretized equations, especially when applied to problems involving curvilinear coordinates encoded in a metric (e.g., in numerical relativity).  
Results from a suite of tests in one and two spatial dimensions --- including problems with strong shocks --- demonstrate reliable performance of the implementation.  
Two tests in spherical symmetry (standing accretion shock \cite{blondin_etal_2003} and Yahil-Lattimer collapse \cite{yahilLattimer_1982,yahil_1983}) demonstrate the DG method's ability to handle conditions relevant to CCSN simulations.  
We will present a more in-depth analysis of the performance of \thornado\ on these (and related) problems in a future study.

@@

\section*{References}
\begin{thebibliography}{34}
  \bibitem{bassi_etal_2013} Bassi F, Franchina N, Ghidoni A and Rebay S 2013 {\it Int. J. Numer. Meth. Fluids} {\bf 71} 1322
  \bibitem{blondin_etal_2003} Blondin J M, Mezzacappa A and DeMarino C 2003 {\it ApJ} {\bf 584} 980
  \bibitem{brennerScott_2008} Brenner S C and Scott L R 2008 {\it The Mathematical Theory of Finite Element Methods} Springer
  \bibitem{cockburnShu_1998} Cockburn B and Shu C-W 1998 {\it JCP} {\bf 141} 199
  \bibitem{cockburnShu_2001} Cockburn B and Shu C-W 2001 {\it Journal of Scientific Computing} {\bf 16} 173
  \bibitem{dumbser_etal_2014} Dumbser M, Zanotti O, Loub{\`e}re R and Diot S 2014 {\it JCP} {\bf 278} 47
  \bibitem{fambri_etal_2018} Fambri F, Dumbser M, K{\"o}ppel S, Rezzolla L and Zanotti O 2018 {\it MNRAS} {\bf 477} 4543
  \bibitem{fuShu_2017} Fu G and Shu C-W 2017 {\it JCP} {\bf 347} 305
  \bibitem{harten_etal_1983} Harten A, Lax P D and Van Leer B 1983 {\it SIAM~Review} {\bf 25} 35
  \bibitem{hesthavenWarburton_2008} Hesthaven J S and Warburton T 2008 {\it Nodal discontinuous Galerkin methods: Algorithms, analysis and applications} Springer
  \bibitem{klockner_etal_2009} Kl{\"o}ckner A, Warburton T, Bridge J and Hesthaven J S 2009 {\it JCP} {\bf 228} 7863
  \bibitem{landauLifshitz_1979} Landau L D and Lifshitz E M 1979 {\it Fluid Mechanics} Pergamon
  \bibitem{liskaWendroff_2003} Liska R and Wendroff B 2003 {\it SIAM J. Sci. Comput.} {\bf 25} 3
  \bibitem{martiMuller_2003} Mart{\'i} J M and M{\"u}ller E 2003 {\it Living Rev. Relativ.} {\bf 6} 7
  \bibitem{mcnally_etal_2012} McNally C, Lyra W and Passy J-C 2012 {\it ApJS} {\bf 201} 18
  \bibitem{mignoneBodo_2005} Mignone A and Bodo G 2005 {\it MNRAS} {\bf 364} 126
  \bibitem{muller_2016} M{\"u}ller B 2016 {\it PASA} {\bf 33} 1
  \bibitem{qin_etal_2016} Qin T, Shu C-W and Yang Y 2016 {\it JCP} {\bf 315} 323
  \bibitem{radiceRezzolla_2011} Radice D and Rezzolla L 2011 {\it Phys. Rev. D} {\bf 84} 024010
  \bibitem{remacle_etal_2003} Remacle J-F, Flaherty J E and Shephard M S 2003 {\it SIAM~Review} {\bf 45} 53
  \bibitem{schaal_etal_2015} Schaal K, Bauer A, Chandrashekar P R{\"u}diger P, Klingenberg C and Springel V 2015 {\it MNRAS} {\bf 453} 4278
  \bibitem{sedov_1959} Sedov L I 1959 {\it Similarity and Dimensional Methods in Mechanics} Academic Press
  \bibitem{shuOsher_1988} Shu C-W and Osher S 1988 {\it JCP} {\bf 77} 439
  \bibitem{shuOsher_1989} Shu C-W and Osher S 1989 {\it JCP} {\bf 83} 32
  \bibitem{shu_2016} Shu C-W 2016 {\it JCP} {\bf 316} 598
  \bibitem{sod_1978} Sod G A 1978 {\it JCP} {\bf 27} 1
  \bibitem{teukolsky_2016} Teukolsky S A 2016 {\it JCP} {\bf 312} 333
  \bibitem{toro_etal_1994} Toro E F, Spruce M and Speares W 1994 {\it Shock Waves} {\bf 4} 25
  \bibitem{toro_1999} Toro E F 1999 {\it NUMERICA: A Library of Source Codes for Teaching, Research and Applications}
  \bibitem{wilson_etal_1996} Wilson J R, Mathews G J and Marronetti P 1996 {\it Phys. Rev. D} {\bf 54} 1317
  \bibitem{wuTang_2015} Wu K and Tang H 2015 {\it JCP} {\bf 298} 539
  \bibitem{yahilLattimer_1982} Yahil A and Lattimer J M 1982 {\it Supernovae: A Survey of Current Research} Reidel
  \bibitem{yahil_1983} Yahil A 1983 {\it ApJ} {\bf 265} 1047
  \bibitem{zhangShu_2010} Zhang X and Shu C-W 2010 {\it JCP} {\bf 229} 8918
\end{thebibliography}

>>>>

\newcommand{\deriv}[2]{\frac{d #1}{d #2}}
\newcommand{\pderiv}[2]{\frac{\partial #1}{\partial #2}}
\newcommand{\vect}[1]{\boldsymbol{#1}}
\newcommand{\f}[2]{\frac{#1}{#2}}
\newcommand{\dx}{\Delta x}
\newcommand{\pd}[2]{\partial_{#2}{#1}}
\newcommand{\bx}{\vect{x}}
\newcommand{\bK}{\vect{K}}
\newcommand{\cT}{\mathcal{T}}
\newcommand{\xL}{x_{\textnormal{\tiny\textsc{L}}}}
\newcommand{\xH}{x_{\textnormal{\tiny\textsc{H}}}}
\newcommand{\xMin}{x_{\textnormal{\tiny m}}}
\newcommand{\xMax}{x_{\textnormal{\tiny M}}}
\newcommand{\trans}{\textnormal{\tiny\textsc{T}}}
\newcommand{\leftState}{\textnormal{\tiny\textsc{L}}}
\newcommand{\rightState}{\textnormal{\tiny\textsc{R}}}
\newcommand{\TCI}{\textnormal{\tiny\textsc{TCI}}}
\newcommand{\shock}{\textnormal{\tiny\textsc{Sh}}}
\newcommand{\thornado}{\texttt{thornado}}
\newcommand{\chimera}{{\sc Chimera}}

@@

We have presented preliminary algorithm details and numerical results for solvers of the non-relativistic and special relativistic Euler equations of gas dynamics as implemented in the {\bf t}oolkit for {\bf h}igh-{\bf or}der {\bf n}eutrino-r{\bf ad}iation hydr{\bf o}dynamics (\thornado).  
The spatial discretization is based on the DG method and the ODEs resulting from this discretization are integrated in time using SSP-RK methods.  
We employ a spectral-type nodal collocation DG approximation based on Legendre-Gauss points for interpolation and numerical quadrature evaluation \cite{bassi_etal_2013}.  
This choice simplifies expressions for the semi-discretized equations, especially when applied to problems involving curvilinear coordinates encoded in a metric (e.g., in numerical relativity).  
Results from a suite of tests in one and two spatial dimensions --- including problems with strong shocks --- demonstrate reliable performance of the implementation.  
The accuracy of \thornado\ on three shock tests (Sod shock tube \cite{sod_1978}, Sedov-Taylor blast wave \cite{sedov_1959}, and Shu-Osher shock tube \cite{shuOsher_1989}) is qualitatively similar to that of the well-established CCSN simulation code \chimera\ \cite{bruenn_etal_2018}, which is based on the PPM finite volume method \cite{colellaWoodward_1984}.  
Two tests in spherical symmetry (standing accretion shock \cite{blondin_etal_2003} and Yahil-Lattimer collapse \cite{yahilLattimer_1982,yahil_1983}) demonstrate the DG method's ability to handle conditions relevant to CCSN simulations.  
We will present a more in-depth analysis of the performance of \thornado\ on these (and related) problems in a future study.

@@

\section*{References}
\begin{thebibliography}{36}
  \bibitem{bassi_etal_2013} Bassi F, Franchina N, Ghidoni A and Rebay S 2013 {\it Int. J. Numer. Meth. Fluids} {\bf 71} 1322
  \bibitem{blondin_etal_2003} Blondin J M, Mezzacappa A and DeMarino C 2003 {\it ApJ} {\bf 584} 980
  \bibitem{brennerScott_2008} Brenner S C and Scott L R 2008 {\it The Mathematical Theory of Finite Element Methods} Springer
  \bibitem{bruenn_etal_2018} Bruenn S W {\it et al} {\it Preprint} arXiv:1809.05608
  \bibitem{cockburnShu_1998} Cockburn B and Shu C-W 1998 {\it JCP} {\bf 141} 199
  \bibitem{cockburnShu_2001} Cockburn B and Shu C-W 2001 {\it Journal of Scientific Computing} {\bf 16} 173
  \bibitem{colellaWoodward_1984} Colella P and Woodward P 1984 {\it JCP} {\bf 54} 174
  \bibitem{dumbser_etal_2014} Dumbser M, Zanotti O, Loub{\`e}re R and Diot S 2014 {\it JCP} {\bf 278} 47
  \bibitem{fambri_etal_2018} Fambri F, Dumbser M, K{\"o}ppel S, Rezzolla L and Zanotti O 2018 {\it MNRAS} {\bf 477} 4543
  \bibitem{fuShu_2017} Fu G and Shu C-W 2017 {\it JCP} {\bf 347} 305
  \bibitem{harten_etal_1983} Harten A, Lax P D and Van Leer B 1983 {\it SIAM~Review} {\bf 25} 35
  \bibitem{hesthavenWarburton_2008} Hesthaven J S and Warburton T 2008 {\it Nodal discontinuous Galerkin methods: Algorithms, analysis and applications} Springer
  \bibitem{klockner_etal_2009} Kl{\"o}ckner A, Warburton T, Bridge J and Hesthaven J S 2009 {\it JCP} {\bf 228} 7863
  \bibitem{landauLifshitz_1979} Landau L D and Lifshitz E M 1979 {\it Fluid Mechanics} Pergamon
  \bibitem{liskaWendroff_2003} Liska R and Wendroff B 2003 {\it SIAM J. Sci. Comput.} {\bf 25} 3
  \bibitem{martiMuller_2003} Mart{\'i} J M and M{\"u}ller E 2003 {\it Living Rev. Relativ.} {\bf 6} 7
  \bibitem{mcnally_etal_2012} McNally C, Lyra W and Passy J-C 2012 {\it ApJS} {\bf 201} 18
  \bibitem{mignoneBodo_2005} Mignone A and Bodo G 2005 {\it MNRAS} {\bf 364} 126
  \bibitem{muller_2016} M{\"u}ller B 2016 {\it PASA} {\bf 33} 1
  \bibitem{qin_etal_2016} Qin T, Shu C-W and Yang Y 2016 {\it JCP} {\bf 315} 323
  \bibitem{radiceRezzolla_2011} Radice D and Rezzolla L 2011 {\it Phys. Rev. D} {\bf 84} 024010
  \bibitem{remacle_etal_2003} Remacle J-F, Flaherty J E and Shephard M S 2003 {\it SIAM~Review} {\bf 45} 53
  \bibitem{schaal_etal_2015} Schaal K, Bauer A, Chandrashekar P R{\"u}diger P, Klingenberg C and Springel V 2015 {\it MNRAS} {\bf 453} 4278
  \bibitem{sedov_1959} Sedov L I 1959 {\it Similarity and Dimensional Methods in Mechanics} Academic Press
  \bibitem{shuOsher_1988} Shu C-W and Osher S 1988 {\it JCP} {\bf 77} 439
  \bibitem{shuOsher_1989} Shu C-W and Osher S 1989 {\it JCP} {\bf 83} 32
  \bibitem{shu_2016} Shu C-W 2016 {\it JCP} {\bf 316} 598
  \bibitem{sod_1978} Sod G A 1978 {\it JCP} {\bf 27} 1
  \bibitem{teukolsky_2016} Teukolsky S A 2016 {\it JCP} {\bf 312} 333
  \bibitem{toro_etal_1994} Toro E F, Spruce M and Speares W 1994 {\it Shock Waves} {\bf 4} 25
  \bibitem{toro_1999} Toro E F 1999 {\it NUMERICA: A Library of Source Codes for Teaching, Research and Applications}
  \bibitem{wilson_etal_1996} Wilson J R, Mathews G J and Marronetti P 1996 {\it Phys. Rev. D} {\bf 54} 1317
  \bibitem{wuTang_2015} Wu K and Tang H 2015 {\it JCP} {\bf 298} 539
  \bibitem{yahilLattimer_1982} Yahil A and Lattimer J M 1982 {\it Supernovae: A Survey of Current Research} Reidel
  \bibitem{yahil_1983} Yahil A 1983 {\it ApJ} {\bf 265} 1047
  \bibitem{zhangShu_2010} Zhang X and Shu C-W 2010 {\it JCP} {\bf 229} 8918
\end{thebibliography}
